Review on precise spiking motifs in neurobiological and neuromorphic data

### 2022-10-04 manuscript.tex
Update manuscript.tex

\subsubsection{\texorpdfstring{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{ref-wVZpQvSk}}{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}42{]}}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{ref-yLVM9IXi}. It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands. Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials. Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events. Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent. These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.

@@

\begin{itemize}
 \item
  https://www.janelia.org/lab/stringer-lab
\item
  described in~\citep{ref-zc16plUu}{100}{]}
\item
  \href{https://github.com/MouseLand/rastermap}{rastermap}
\item
  deconvolution strategy
\item
  based on a linear model
\end{itemize}

\subsubsection{\texorpdfstring{Stringer et al 2019, Nature~\citep{ref-brg2FFsb}{101}{]}}{Stringer et al 2019, Nature {[}101{]}}}\label{stringer-et-al-2019-nature-stringer2019nature}
\begin{itemize}
 \item
  ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated. Here we analysed the dimensionality of the encoding of natural images by large populations of neurons in the visual cortex of awake mice.''
\item
  Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
\item
  Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
\end{itemize}

\subsubsection{\texorpdfstring{Stringer et al 2019, Science~\citep{ref-1BPHUQLTf}{102}{]}}{Stringer et al 2019, Science {[}102{]}}}\label{stringer-et-al-2019-science-stringer2019science}
\begin{itemize}
\item
  Stringer et al.~analyzed spontaneous neuronal firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements. The variability of neuronal responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
\item
  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{ref-1EMs3GS0B}{103}{]}
\end{itemize}

\subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}
\subsubsection{\texorpdfstring{Paper by Grossberger, 2018~\citep{ref-9AK1cwY1}{104}{]}}{Paper by Grossberger, 2018 {[}104{]}}}\label{paper-by-grossberger-2018-grossberger2018}
\begin{itemize}
 \item
  Temporally ordered multi-neuron patterns likely encode information in the brain. We introduce an unsupervised method, SPOTDisClust (Spike Pattern Optimal Transport Dissimilarity Clustering), for their detection from high-dimensional neural ensembles. SPOTDisClust measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other (SPOTDis).
\item
  Detecting these temporal patterns represents a major methodological challenge.
\item
  Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
\item
  Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
\item
  Lee AK, Wilson MA. A combinatorial method for analyzing sequential firing patterns involving an arbitrary number of neurons based on relative time order. J Neurophysiol. 2004;92(4):2555--73. pmid:15212425
\item
  Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
\item
  only one spike per neuron: fig 1A = ``For each pattern and each neuron, a random position was chosen for the activation pulse.''
\item
  t-SNE projection with HDBSCAN labels shows that our clustering method can retrieve all patterns from the data.
\item
  data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
\end{itemize}

\begin{figure}
\centering
\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
\caption{Fig 1 of~\citep{ref-9AK1cwY1}{104}{]}: ``Simulated example illustrating the steps in SPOTDisClust. A) Structure of five ``ground-truth'' patterns (\ldots). For each pattern and each neuron, a random position was chosen for the activation pulse. B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A)." (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:G2018-1}
\end{figure}

@@

\subsubsection{\texorpdfstring{Paper by Russo et al 2017~\citep{ref-RrgzWOSk}}{Paper by Russo et al 2017 {[}88{]}}}\label{paper-by-russo-et-al-2017-russo2017}
\begin{itemize}
 \item
  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{ref-RrgzWOSk}
\item
  sliding window as in~\citep{ref-1XpTamcl}{105}{]} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{ref-bc9czTxi}
\item
  based on a ``non-stationarity-corrected parametric test statistic for assessing the independence of pairs'' and ``an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies''
\end{itemize}

\subsubsection{\texorpdfstring{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{ref-1300cS9RL}{106}{]}}{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex {[}106{]}}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
\begin{itemize}
 \item
  Stochastic sampling links perceptual uncertainty to neural response variability
\item
  Model accounts for independent changes in strength and variability of responses
\item
  Model predicts relationship between noise, signal, and spontaneous correlations
\item
  Stimulus statistics dependence of response statistics is explained
\end{itemize}

To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{ref-1365b3Wwl}{107}{]} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time. They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.

\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}\label{dynamics-of-delay-coupled-excitable-neural-systems}
uses a FPGA
~\citep{ref-rYLmAESf}{108}{]}

@@

V. Thanasoulis, B. Vogginger, J. Partzsch and C. Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
~\citep{ref-1DadetlqZ}{109}{]}

\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
~\citep{ref-RrgzWOSk} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence. It is based on unitary events in multiple single-neuron spiking activity~\citep{ref-iNkKTYaI}{110}{]} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{ref-14wpaupd5}{111}{]}. Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.

\section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}
\subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{ref-19NyVq80B}{112}{]} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays. This STDP rule will be based by the inversion of the generative model for spike formation and will therefore be derived by a Bayesian approach. This will decouple the active synapses (similarly to a logistic regression) from the values of possible synaptic delays.

\begin{itemize}
 \item
 ~\citep{ref-11p3NXmkg}{113}{]} : coherence detection
\item
 ~\citep{ref-j5xS3owo}{114}{]} : STDP
\item
  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{ref-15TktA0gR}{115}{]} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs. PLoS Comput Biol 9(2): e1002897.~\citep{ref-1FHIKpizQ}{116}{]} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
\end{itemize}

Our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision. This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{ref-s09KZ8Kh}{117}{]}

@@

~\citep{ref-UeTLXQCp}{118}{]}

@@

In a recent paper~\citep{ref-B493Fl4K}{119}{]}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron. In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.

In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{ref-pPRQwZb}{120}{]}. This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.

\subsection{Learning sequences}\label{learning-sequences}
In~\citep{ref-WFwUKZb5}{121}{]}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga}{124}{]} and this model shows how one could vary HVC activity using something similar to dropout in ML. Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).

\begin{itemize}
\item
  in~\citep{ref-Jzr1gwsr}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
\item
  In Bellec~\citep{ref-aj6TgorR}{125}{]}, authors fit summary statistics of neural data with a differentiable spiking network simulator.

  \begin{itemize}
     \item
    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{ref-14e3BCBii}{126}{]} with recurrent dynamics)
  \item
    sample and measure method to include latent / hidden neurons
  \item
    comes with code https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure
  \item
    V1-dataset : The dataset we used was collected by Smith and Kohn {[}49{]} and is publicly available at:
    http://crcns.org/data-sets/vc/pvc-11 - it is in a sense supervised with the input being the movie and the output the spikes recorded.
  \end{itemize}
\end{itemize}

A recent work~\citep{ref-10EU8MfxC}{127}{]} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream. The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern. The second stage takes the output of the spatiotemporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatiotemporal activity. For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work. Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters. Showing the potential of such representation

@@

Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
Pang R, Fairhall AL.Elife. 2019 May~\citep{ref-12zyUFdO4}{128}{]}

@@

\subsection{dynamical models}\label{dynamical-models}
Dumas and colleagues~\citep{ref-WxQdcMB2}{129}{]} : three levels / fourth paradigm~\citep{ref-JrZPoxk9}{130}{]} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.

@@

Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{ref-sJ03d2Zj}{131}{]} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}{132}{]}.

@@

\section{development}\label{development}
scaffolding of neural assemblies / existence of critical periods :~\citep{ref-1HSyNOl48}{133}{]}

>>>>

\subsubsection{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function~\citep{ref-wVZpQvSk}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}
It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas~\citep{ref-yLVM9IXi}. It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands. Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials. Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events. Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent. These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.

@@

\begin{itemize}
 \item
  https://www.janelia.org/lab/stringer-lab
\item
  described in~\citep{ref-zc16plUu}
\item
  \href{https://github.com/MouseLand/rastermap}{rastermap}
\item
  deconvolution strategy
\item
  based on a linear model
\end{itemize}

\subsubsection{Stringer et al 2019, Nature~\citep{ref-brg2FFsb}}\label{stringer-et-al-2019-nature-stringer2019nature}
\begin{itemize}
 \item
  ``A neuronal population encodes information most efficiently when its stimulus responses are high-dimensional and uncorrelated, and most robustly when they are lower-dimensional and correlated. Here we analysed the dimensionality of the encoding of natural images by large populations of neurons in the visual cortex of awake mice.''
\item
  Data availability: All of the processed deconvolved calcium traces are available on \href{https://figshare.com/articles/Recordings_of_ten_thousand_neurons_in_visual_cortex_in_response_to_2_800_natural_images/6845348}{figshare}, together with the image stimuli.
\item
  Code availability: The code is available on \href{https://github.com/MouseLand/stringer-pachitariu-et-al-2018b}{GitHub}.
\end{itemize}

\subsubsection{Stringer et al 2019, Science~\citep{ref-1BPHUQLTf}}\label{stringer-et-al-2019-science-stringer2019science}
\begin{itemize}
\item
  Stringer et al.~analyzed spontaneous neuronal firing, finding that neurons in the primary visual cortex encoded both visual information and motor activity related to facial movements. The variability of neuronal responses to visual stimuli in the primary visual area is mainly related to arousal and reflects the encoding of latent behavioral states.
\item
  see also the work showing that you can encode very precise orientation information by using many neurons:~\citep{ref-1EMs3GS0B}
\end{itemize}

\subsection{detecting structured temporal patterns}\label{detecting-structured-temporal-patterns}
\subsubsection{Paper by Grossberger, 2018~\citep{ref-9AK1cwY1}}\label{paper-by-grossberger-2018-grossberger2018}
\begin{itemize}
 \item
  Temporally ordered multi-neuron patterns likely encode information in the brain. We introduce an unsupervised method, SPOTDisClust (Spike Pattern Optimal Transport Dissimilarity Clustering), for their detection from high-dimensional neural ensembles. SPOTDisClust measures similarity between two ensemble spike patterns by determining the minimum transport cost of transforming their corresponding normalized cross-correlation matrices into each other (SPOTDis).
\item
  Detecting these temporal patterns represents a major methodological challenge.
\item
  Many approaches to this problem are supervised, that is, they take patterns occurring concurrently with a known event, such as the delivery of a stimulus for sensory neurons or the traversal of a running track for hippocampal place fields, as a ``template'' and then search for repetitions of the same template in spiking activity :
\item
  Nadasdy Z, Hirase H, Czurko A, Csicsvari J, Buzsaki G. Replay and time compression of recurring spike sequences in the hippocampus. J Neurosci. 1999;19(21):9497--507. pmid:10531452
\item
  Lee AK, Wilson MA. A combinatorial method for analyzing sequential firing patterns involving an arbitrary number of neurons based on relative time order. J Neurophysiol. 2004;92(4):2555--73. pmid:15212425
\item
  Davidson TJ, Kloosterman F, Wilson MA. Hippocampal replay of extended experience. Neuron. 2009;63(4):497--507. pmid:19709631
\item
  only one spike per neuron: fig 1A = ``For each pattern and each neuron, a random position was chosen for the activation pulse.''
\item
  t-SNE projection with HDBSCAN labels shows that our clustering method can retrieve all patterns from the data.
\item
  data available @ https://doi.org/10.1371/journal.pcbi.1006283.s013
\end{itemize}

\begin{figure}
\centering
\includegraphics[width=7in,height=\textheight]{images/pcbi.1006283.g001.PNG_L.png}
\caption{Fig 1 of~\citep{ref-9AK1cwY1}: ``Simulated example illustrating the steps in SPOTDisClust. A) Structure of five ``ground-truth'' patterns (\ldots). For each pattern and each neuron, a random position was chosen for the activation pulse. B) Neuronal output is generated according to an inhomogeneous Poisson process, with rates dictated by the patterns in (A)." (© Authors under a \href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006283}{CC licence})}\label{fig:G2018-1}
\end{figure}

@@

\subsubsection{Paper by Russo et al 2017~\citep{ref-RrgzWOSk}}\label{paper-by-russo-et-al-2017-russo2017}
\begin{itemize}
 \item
  ``Here we present such a unifying methodological and conceptual framework which detects assembly structure at many different time scales, levels of precision, and with arbitrary internal organization.'' by~\citep{ref-RrgzWOSk}
\item
  sliding window as in~\citep{ref-1XpTamcl} (``Numerous other statistical procedures for detecting assemblies or sequential patterns have been proposed previously'') - extended to multiple lags~\citep{ref-bc9czTxi}
\item
  based on a ``non-stationarity-corrected parametric test statistic for assessing the independence of pairs'' and ``an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies''
\end{itemize}

\subsubsection{Neural Variability and Sampling-Based Probabilistic Representations in the Visual Cortex~\citep{ref-1300cS9RL}}\label{neural-variability-and-sampling-based-probabilistic-representations-in-the-visual-cortex-orban2016}
\begin{itemize}
 \item
  Stochastic sampling links perceptual uncertainty to neural response variability
\item
  Model accounts for independent changes in strength and variability of responses
\item
  Model predicts relationship between noise, signal, and spontaneous correlations
\item
  Stimulus statistics dependence of response statistics is explained
\end{itemize}

To overcome the limits of models which require spike times to be discretized, utilize a sub-optimal least-squares criterion, or do not provide uncertainty estimates for model predictions or estimated parameters,~\citep{ref-1365b3Wwl} address each of these shortcomings by developing a point process model that characterizes fine-scale sequences at the level of individual spikes and represents sequence occurrences as a small number of marked events in continuous time. They also introduce learnable time warping parameters to model sequences of varying duration, which have been experimentally observed in neural circuits and demonstrate these advantages on experimental recordings from songbird higher vocal center and rodent hippocampus.

\subsubsection{Dynamics of Delay-Coupled Excitable Neural Systems}\label{dynamics-of-delay-coupled-excitable-neural-systems}
uses a FPGA
~\citep{ref-rYLmAESf}

@@

V. Thanasoulis, B. Vogginger, J. Partzsch and C. Mayr, ``Delay-Based Neural Computation: Pulse Routing Architecture and Benchmark Application in FPGA,'' 2021 28th IEEE International Conference on Electronics, Circuits, and Systems (ICECS), 2021, pp.~1-5
~\citep{ref-1DadetlqZ}

\subsubsection{Cell assemblies at multiple time scales with arbitrary lag constellations}\label{cell-assemblies-at-multiple-time-scales-with-arbitrary-lag-constellations}
~\citep{ref-RrgzWOSk} starts from the relatively old notion of assessing the departure of the joint spike count distribution of two units (or sets) from independence. It is based on unitary events in multiple single-neuron spiking activity~\citep{ref-iNkKTYaI} and the reliable and efficient analysis of an excess or deficiency of joint-spike events~\citep{ref-14wpaupd5}. Having derived a fast, non-stationarity-corrected parametric test statistic for assessing the independence of pairs, they designed an agglomerative, heuristic clustering algorithm for fusing significant pairs into higher-order assemblies.

\section{Learning to detect polychronous groups}\label{learning-to-detect-polychronous-groups}
\subsection{Learning weights \ldots{} and delays}\label{learning-weights-and-delays}
spike time coding in a neuron: We will describe the Spike-Time Dependent Plasticity (STDP)~\citep{ref-19NyVq80B} rule which implement an unsupervised learning aiming at optimizing the detection of polychronous patterns, that is volleys of spikes which are synchronized, up to some stable pattern of pre-synaptic delays. This STDP rule will be based by the inversion of the generative model for spike formation and will therefore be derived by a Bayesian approach. This will decouple the active synapses (similarly to a logistic regression) from the values of possible synaptic delays.

\begin{itemize}
 \item
 ~\citep{ref-11p3NXmkg} : coherence detection
\item
 ~\citep{ref-j5xS3owo} : STDP
\item
  Masquelier, Gilson et al.~- 2010 - STDP in Recurrent Neuronal Networks~\citep{ref-15TktA0gR} or Delay Selection by Spike-Timing-Dependent Plasticity in Recurrent Networks of Spiking Neurons Receiving Oscillatory Inputs. PLoS Comput Biol 9(2): e1002897.~\citep{ref-1FHIKpizQ} which targets at the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity.
\end{itemize}

Our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision. This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment.~\citep{ref-s09KZ8Kh}

@@

~\citep{ref-UeTLXQCp}

@@

In a recent paper~\citep{ref-B493Fl4K}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of single spiking neuron. In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.

In a computational model, authors show that the frequently activated polychronous neuronal groups can be learned by readout neurons with joint weight-delay spike-timing-dependent plasticity~\citep{ref-pPRQwZb}. This uses a joint Weight-Delay Spike-Timing-Dependent Plasticity and show aan efficient learning of polychronous groups.

\subsection{Learning sequences}\label{learning-sequences}
In~\citep{ref-WFwUKZb5}, authors present a model to ``show a way by which the nervous system maintains precise, stereotyped behavior in the face of environmental and neural changes.'' It is shown in bridsong generation that ``A precise, temporally sparse sequence from the premotor nucleus HVC is crucial to the performance of song in songbirds''~\citep{ref-1EhOfHrje}{122},\protect\hyperlink{ref-Dry1ANMD}{123},\protect\hyperlink{ref-SAIh1pga} and this model shows how one could vary HVC activity using something similar to dropout in ML. Using such controlled variability, ``behaviors are made more robust to environmental change by continually seeking subtly new ways of performing the same task.'' Not sure however how important it is that the HVC pattern should be sparse (and similar to PGs).

\begin{itemize}
\item
  in~\citep{ref-Jzr1gwsr}, there are ``good'' and ``bad'' noises show that some patterns are more easy to disentangle - similar to bird songs and ecological niche.
\item
  In Bellec~\citep{ref-aj6TgorR}, authors fit summary statistics of neural data with a differentiable spiking network simulator.

  \begin{itemize}
     \item
    the loss function is the cross entropy (following Bernouilli hypothesis with a GLM where each unit is modelled with a SRM neuron~\citep{ref-14e3BCBii} with recurrent dynamics)
  \item
    sample and measure method to include latent / hidden neurons
  \item
    comes with code https://github.com/EPFL-LCN/pub-bellec-wang-2021-sample-and-measure
  \item
    V1-dataset : The dataset we used was collected by Smith and Kohn {[}49{]} and is publicly available at:
    http://crcns.org/data-sets/vc/pvc-11 - it is in a sense supervised with the input being the movie and the output the spikes recorded.
  \end{itemize}
\end{itemize}

A recent work~\citep{ref-10EU8MfxC} proposes a two-stage unsupervised-supervised system for the categorization of spatiotemporal actions from an event-based stream. The first stage learns spatiotemporal convolutional filters targeted to minimize event-removal related changes to a local spatiotemporal spike-event pattern. The second stage takes the output of the spatiotemporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatiotemporal activity. For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work. Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters. Showing the potential of such representation

@@

Fast and flexible sequence induction in spiking neural networks via rapid excitability changes.
Pang R, Fairhall AL.Elife. 2019 May~\citep{ref-12zyUFdO4}

@@

\subsection{dynamical models}\label{dynamical-models}
Dumas and colleagues~\citep{ref-WxQdcMB2} : three levels / fourth paradigm~\citep{ref-JrZPoxk9} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.

@@

Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{ref-sJ03d2Zj} or of sparsity aware algorithms~\citep{ref-R5BYFgCP}.

@@

\section{development}\label{development}
scaffolding of neural assemblies / existence of critical periods :~\citep{ref-1HSyNOl48}
